---
title: "2.SFS_contemporary_intergenic_X_vs_autosomic"
output: html_document
---

# Copy transformed thetas from datos.

All my "$POP".transformedThetas are already in datos:

/home/mlucena/datos/intermediate_ANGSD_files/intergenic_analysis/intergenic_sfs

Therefore, I will copy them:

```{bash}
screen -S copying_thetas_file

scp /home/mlucena/datos/intermediate_ANGSD_files/intergenic_analysis/intergenic_sfs/*transformedThetas /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs

```

# XCHR

## Create folder for SFS diversity for X chr intergenic and autosomic

I will create a folder where I will save the cut thetas file. 
```{bash}

mkdir /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_Xchr/
cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_Xchr/

```


## Create folders separated by scaffolds for diversity calculus per unit 

Mi rf file está en:

/GRUPOS/grupolince/Lynx_Xchr/Xchr.intergenic.bed

Para este analisis tenemos que rompar el bed file con el que queremos hacer el intersect en distintos scaffolds, porque no puede cargar el archivo de diversidades (thetas )
para hacer un intersect con todo el bed.Así que vamos a partirlo en scaffold y despues buscar para cada scaffolds que regiones quiero y así costruir un archivo 
thetas de diversidades de X intergénico. 


```{bash}

cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_Xchr

screen -S folder_for_diversity_for_chrX_analysis

scp /GRUPOS/grupolince/Lynx_Xchr/Xchr.intergenic.bed /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_Xchr

TARGETBED=/home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_Xchr/Xchr.intergenic.bed
TARGET_FOLDER=$TARGETBED.PerScaffold

mkdir $TARGET_FOLDER
cd $TARGET_FOLDER

scp $TARGETBED $TARGET_FOLDER

#Split per scaffold
# Create multiple files base on one column: transformedThetas:
awk '{print >> $1; close($1)}' "$TARGETBED"

# Rename the files: splitted_annotations
for file in lp23*
do
echo $file
mv $file ${file/lp23/Xchr.intergenic.bed_lp23}
done

cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_Xchr/Xchr.intergenic.bed.PerScaffold
rm Xchr.intergenic.bed

```

# Autosomic

First I will create the folder:
```{bash}
mkdir /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_autosomic
cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_autosomic
```

## Create folders separated by scaffolds for diversity calculus per unit 

Mi rf file está en:

/home/mlucena/grupolince/reference_genomes/lynx_pardinus_genome/bed_file_all_the_genome_without_Y_X_chr.bed

Igual que antes rompemos el archivo en los distintos cromosomas:
```{bash}

cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_autosomic

screen -S folder_for_diversity_for_autosomic_analysis

scp /home/mlucena/grupolince/reference_genomes/lynx_pardinus_genome/bed_file_all_the_genome_without_Y_X_chr.bed /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_autosomic

TARGETBED=/home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_autosomic/bed_file_all_the_genome_without_Y_X_chr.bed
TARGET_FOLDER=$TARGETBED.PerScaffold

mkdir $TARGET_FOLDER
cd $TARGET_FOLDER

scp $TARGETBED $TARGET_FOLDER

#Split per scaffold
# Create multiple files base on one column: transformedThetas:
awk '{print >> $1; close($1)}' "$TARGETBED"

# Rename the files: splitted_annotations
for file in lp23*
do
echo $file
mv $file ${file/lp23/autosomic.intergenic.bed_lp23}
done

cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_autosomic/bed_file_all_the_genome_without_Y_X_chr.bed.PerScaffold
rm bed_file_all_the_genome_without_Y_X_chr.bed

```


# Diversity Calculus

## Defining variables

I checked that all bam files were indexed --> OK

##Create script file

To run in loop

###Open nano

```{bash}
cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs
nano  script_from_saf_to_diversity_Xchr_vs_autosomic.sh
```


El archivo donde se almacenan las diversidades lo tengo que crear antes.

```{bash}
# Create headers for the outfile
echo -e "pop\tregion\tpositions\twatterson_ave\twatterson_sd\tpairwise_ave\tpairwise_sd\ttajimaD" > diversity.per.pop.averages_Xchr.tsv

# Create headers for the outfile
echo -e "pop\tregion\tpositions\twatterson_ave\twatterson_sd\tpairwise_ave\tpairwise_sd\ttajimaD" > diversity.per.pop.averages_autosomic.tsv

```



###Script itself

```{bash}
#!/bin/bash
POP=$(cat /dev/stdin)


##Split transformedThetas per scaffold.

DIR_thetafile=/home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs
cd "$DIR_thetafile"
# mkdir "$DIR_thetafile"/diversity_per_scaffold/
mkdir "$DIR_thetafile"/diversity_per_scaffold/"$POP"_separated_by_scaffold/
mv "$POP".transformedThetas "$DIR_thetafile"/diversity_per_scaffold/"$POP"_separated_by_scaffold/
cd  "$DIR_thetafile"/diversity_per_scaffold/"$POP"_separated_by_scaffold/
# Create multiple files base on one column: transformedThetas:
awk '{print >> $1; close($1)}' "$POP".transformedThetas
rm scaffold ### no se pq esto? porque aparece un archivo, el de la primera línea, que se llama scaffold

# Rename the files: transformedThetas
for file in lp23*
do
echo $file
mv $file ${file/lp23/"$POP".transformedThetas_lp23}
done

##Diversity x Chomosome.

## WG variables

cd "$DIR_thetafile"/diversity_per_scaffold/"$POP"_separated_by_scaffold/
SCAFFOLDS_FOLDER_X=/home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_Xchr/Xchr.intergenic.bed.PerScaffold/
cd $SCAFFOLDS_FOLDER_X
SCAFFOLDS=($(find $SCAFFOLDS_FOLDER_X -name "Xchr.intergenic.bed*" -exec ls {} \; | awk '{ print substr( $0, length($0) - 10, length($0) ) }'  | LANG=en_EN sort | uniq ))
DIR=/home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_Xchr/
SFS_FOLDER="$DIR"
ANALYIS="X_chr"

cd $DIR

for SCAFFOLD in ${SCAFFOLDS[@]}
do
echo "---------------------$SCAFFOLD---------------------"
bedtools intersect -a <(cat "$DIR_thetafile"/diversity_per_scaffold/"$POP"_separated_by_scaffold/"$POP".transformedThetas_"$SCAFFOLD") -b <(cat "$SCAFFOLDS_FOLDER_X"Xchr.intergenic.bed_"$SCAFFOLD") >> "$POP".transformedThetas_Xchr
done




WATTERSON_AVERAGE=$(cut -f 4 "$POP".transformedThetas_Xchr | awk -v OFS='\t' '{for(i=1;i<=NF;i++) {sum[i] += $i }} END {for (i=1;i<=NF;i++) {printf ("%.10e\n", sum[i]/NR )}}' | sed 's/[eE]+\{0,1\}/*10^/g')
WATTERSON_SD=$(cut -f 4 "$POP".transformedThetas_Xchr | awk -v OFS='\t' '{for(i=1;i<=NF;i++) {sum[i] += $i; sumsq[i] += ($i)^2}} END {for (i=1;i<=NF;i++) {printf ("%.10e\n", sqrt((sumsq[i]-sum[i]^2/NR)/NR))}}' | sed 's/[eE]+\{0,1\}/*10^/g')
PAIRWISE_AVERAGE=$(cut -f 5 "$POP".transformedThetas_Xchr | awk -v OFS='\t' '{for(i=1;i<=NF;i++) {sum[i] += $i }} END {for (i=1;i<=NF;i++) {printf ("%.10e\n", sum[i]/NR )}}' | sed 's/[eE]+\{0,1\}/*10^/g') 
PAIRWISE_SD=$(cut -f 5 "$POP".transformedThetas_Xchr | awk -v OFS='\t' '{for(i=1;i<=NF;i++) {sum[i] += $i; sumsq[i] += ($i)^2}} END {for (i=1;i<=NF;i++) {printf ("%.10e\n", sqrt((sumsq[i]-sum[i]^2/NR)/NR))}}' | sed 's/[eE]+\{0,1\}/*10^/g') 
DIFFERENCE_PAIRWISE_WATTERSON_SD=$(cut -f 6  "$POP".transformedThetas_Xchr |  awk -v OFS='\t' '{for(i=1;i<=NF;i++) {sum[i] += $i; sumsq[i] += ($i)^2}} END {for (i=1;i<=NF;i++) {printf ("%.10e\n", sqrt((sumsq[i]-sum[i]^2/NR)/NR ))}}' | sed 's/[eE]+\{0,1\}/*10^/g') 
TAJIMAS_D=$(echo "(($PAIRWISE_AVERAGE) - ($WATTERSON_AVERAGE))/($DIFFERENCE_PAIRWISE_WATTERSON_SD)" | bc -l | awk '{printf ("%.10e\n",$1)}' |  sed 's/[eE]+\{0,1\}/*10^/g' )
NR=$(wc -l  "$POP".transformedThetas_Xchr  | cut -d" " -f1)


#Before printing check all variables are full, if empty use NA
if [ -z ${POP} ]; then  POP=NA;  fi
if [ -z ${NR} ]; then  NR=NA;  fi
if [ -z ${WATTERSON_AVERAGE} ]; then  WATTERSON_AVERAGE=NA;  fi
if [ -z ${WATTERSON_SD} ]; then  WATTERSON_SD=NA;  fi
if [ -z ${PAIRWISE_AVERAGE} ]; then  PAIRWISE_AVERAGE=NA;  fi
if [ -z ${PAIRWISE_SD} ]; then  PAIRWISE_SD=NA;  fi
if [ -z ${DIFFERENCE_PAIRWISE_WATTERSON_SD} ]; then  DIFFERENCE_PAIRWISE_WATTERSON_SD=NA;  fi
if [ -z ${TAJIMAS_D} ]; then  TAJIMAS_D=NA;  fi
if [ -z ${ANALYIS} ]; then  ANALYIS=NA;  fi



paste \
<(echo $POP) \
<(echo $ANALYIS) \
<(echo $NR) \
<(echo $WATTERSON_AVERAGE) \
<(echo $WATTERSON_SD) \
<(echo $PAIRWISE_AVERAGE) \
<(echo $PAIRWISE_SD) \
<(echo $TAJIMAS_D) |\
sed 's/ /\t/g'| sed 's/\t\+/\t/g' >>  diversity.per.pop.averages_Xchr.tsv


unset NR
unset ANALYIS
unset WATTERSON_AVERAGE
unset WATTERSON_SD
unset PAIRWISE_AVERAGE
unset PAIRWISE_SD
unset TAJIMAS_D
unset DIFFERENCE_PAIRWISE_WATTERSON_SD
unset SCAFFOLDS
unset DIR


###################################################################
##Diversity autosomic

## WG variables

cd "$DIR_thetafile"/diversity_per_scaffold/"$POP"_separated_by_scaffold/
SCAFFOLDS_FOLDER_autosomic=/home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_autosomic/bed_file_all_the_genome_without_Y_X_chr.bed.PerScaffold/
cd $SCAFFOLDS_FOLDER_autosomic
SCAFFOLDS=($(find $SCAFFOLDS_FOLDER_autosomic -name "autosomic.intergenic.bed*" -exec ls {} \; | awk '{ print substr( $0, length($0) - 10, length($0) ) }'  | LANG=en_EN sort | uniq ))
DIR=/home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_autosomic
SFS_FOLDER="$DIR"
ANALYIS="autosomic"

cd $DIR

for SCAFFOLD in ${SCAFFOLDS[@]}
do
echo "---------------------$SCAFFOLD---------------------"
bedtools intersect -a <(cat "$DIR_thetafile"/diversity_per_scaffold/"$POP"_separated_by_scaffold/"$POP".transformedThetas_"$SCAFFOLD") -b <(cat "$SCAFFOLDS_FOLDER_autosomic"autosomic.intergenic.bed_"$SCAFFOLD") >> "$POP".transformedThetas_autosomic
done



WATTERSON_AVERAGE=$(cut -f 4 "$POP".transformedThetas_autosomic | awk -v OFS='\t' '{for(i=1;i<=NF;i++) {sum[i] += $i }} END {for (i=1;i<=NF;i++) {printf ("%.10e\n", sum[i]/NR )}}' | sed 's/[eE]+\{0,1\}/*10^/g')
WATTERSON_SD=$(cut -f 4 "$POP".transformedThetas_autosomic | awk -v OFS='\t' '{for(i=1;i<=NF;i++) {sum[i] += $i; sumsq[i] += ($i)^2}} END {for (i=1;i<=NF;i++) {printf ("%.10e\n", sqrt((sumsq[i]-sum[i]^2/NR)/NR))}}' | sed 's/[eE]+\{0,1\}/*10^/g')
PAIRWISE_AVERAGE=$(cut -f 5 "$POP".transformedThetas_autosomic | awk -v OFS='\t' '{for(i=1;i<=NF;i++) {sum[i] += $i }} END {for (i=1;i<=NF;i++) {printf ("%.10e\n", sum[i]/NR )}}' | sed 's/[eE]+\{0,1\}/*10^/g') 
PAIRWISE_SD=$(cut -f 5 "$POP".transformedThetas_autosomic | awk -v OFS='\t' '{for(i=1;i<=NF;i++) {sum[i] += $i; sumsq[i] += ($i)^2}} END {for (i=1;i<=NF;i++) {printf ("%.10e\n", sqrt((sumsq[i]-sum[i]^2/NR)/NR))}}' | sed 's/[eE]+\{0,1\}/*10^/g') 
DIFFERENCE_PAIRWISE_WATTERSON_SD=$(cut -f 6  "$POP".transformedThetas_autosomic |  awk -v OFS='\t' '{for(i=1;i<=NF;i++) {sum[i] += $i; sumsq[i] += ($i)^2}} END {for (i=1;i<=NF;i++) {printf ("%.10e\n", sqrt((sumsq[i]-sum[i]^2/NR)/NR ))}}' | sed 's/[eE]+\{0,1\}/*10^/g') 
TAJIMAS_D=$(echo "(($PAIRWISE_AVERAGE) - ($WATTERSON_AVERAGE))/($DIFFERENCE_PAIRWISE_WATTERSON_SD)" | bc -l | awk '{printf ("%.10e\n",$1)}' |  sed 's/[eE]+\{0,1\}/*10^/g' )
NR=$(wc -l  "$POP".transformedThetas_autosomic  | cut -d" " -f1)


#Before printing check all variables are full, if empty use NA
if [ -z ${POP} ]; then  POP=NA;  fi
if [ -z ${NR} ]; then  NR=NA;  fi
if [ -z ${WATTERSON_AVERAGE} ]; then  WATTERSON_AVERAGE=NA;  fi
if [ -z ${WATTERSON_SD} ]; then  WATTERSON_SD=NA;  fi
if [ -z ${PAIRWISE_AVERAGE} ]; then  PAIRWISE_AVERAGE=NA;  fi
if [ -z ${PAIRWISE_SD} ]; then  PAIRWISE_SD=NA;  fi
if [ -z ${DIFFERENCE_PAIRWISE_WATTERSON_SD} ]; then  DIFFERENCE_PAIRWISE_WATTERSON_SD=NA;  fi
if [ -z ${TAJIMAS_D} ]; then  TAJIMAS_D=NA;  fi
if [ -z ${ANALYIS} ]; then  ANALYIS=NA;  fi



paste \
<(echo $POP) \
<(echo $ANALYIS) \
<(echo $NR) \
<(echo $WATTERSON_AVERAGE) \
<(echo $WATTERSON_SD) \
<(echo $PAIRWISE_AVERAGE) \
<(echo $PAIRWISE_SD) \
<(echo $TAJIMAS_D) |\
sed 's/ /\t/g'| sed 's/\t\+/\t/g' >>  diversity.per.pop.averages_autosomic.tsv


unset NR
unset ANALYIS
unset WATTERSON_AVERAGE
unset WATTERSON_SD
unset PAIRWISE_AVERAGE
unset PAIRWISE_SD
unset TAJIMAS_D
unset DIFFERENCE_PAIRWISE_WATTERSON_SD
unset SCAFFOLDS
unset DIR

# rm -r "$DIR_thetafile"/diversity_per_scaffold/"$POP"_separated_by_scaffold

```

###Chmod

```{bash}

chmod 777 script_from_saf_to_diversity_Xchr_vs_autosomic.sh
chmod 777 script_from_saf_to_diversity_Xchr_vs_autosomic_small.sh

```


###Run loop

```{r, engine=bash, eval=FALSE}

cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs

POPS=$(ls *.transformedThetas | cut -d"." -f 1 | uniq )

# Cuidado con el script que quieres lanzar

for POP in ${POPS[@]}
do
echo $POP
nohup echo $POP | ./script_from_saf_to_diversity_Xchr_vs_autosomic.sh > $POP.log 2>&1&
echo $! > $POP_save_pid.txt
done


###
#El script global no ha funcionado para autosomi, asi que lo lanzo solo despuest

cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs

POPS=$(ls *.transformedThetas | cut -d"." -f 1 | uniq )


for POP in ${POPS[@]}
do
echo $POP
nohup echo $POP | ./script_autosomic.sh > $POP.log 2>&1&
echo $! > $POP_save_pid.txt
done


```


## R representation

First I contatenate the tables.

```{bash}
# First I create a directory:
# mkdir /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_comparison_X_autosomic

cat /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_autosomic/diversity.per.pop.averages_autosomic.tsv /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_intergenic_Xchr/diversity.per.pop.averages_Xchr.tsv > /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_comparison_X_autosomic/diversity.per.pop.averages_autosomic_vs_Xchr.tsv

# I download this folder because it is easy to run R in local. 

scp mlucena@genomics-b.ebd.csic.es:/home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_comparison_X_autosomic/diversity.per.pop.averages_autosomic_vs_Xchr.tsv /Users/marialucenaperez/Owncloud/publico/Eurasian_lynx_phylogeography/ANGSD_analysis/analysis_sfs

```

```{r}

library(dplyr) 
library(RColorBrewer)
library(ggplot2)
library(tidyr)

wd <- "/Users/marialucenaperez/Owncloud/publico/Eurasian_lynx_phylogeography/ANGSD_analysis/analysis_sfs/" 
diversity <- read.csv(paste(wd, "diversity.per.pop.averages_autosomic_vs_Xchr.tsv", sep =""), sep ="\t", dec=".", header=T) %>%  
 mutate (., Populations =  ifelse (pop == "c_ll_po_n008", "Bialowieza",
                           ifelse (pop == "c_ll_ur_n006", "Urals", 
                           ifelse (pop == "c_ll_ki_n013", "Kirov", 
                           ifelse (pop == "c_ll_la_n006", "Latvia", 
                           ifelse (pop == "c_ll_no_n008", "Norway", 
                           ifelse (pop == "x_ll_ba_n003", "Balkans" ,
                           ifelse (pop == "c_ll_to_n002", "Töv",
                           ifelse (pop == "c_ll_ka_n004", "Khentii", 
                           ifelse (pop == "c_ll_og_n002", "Ömnögovi", 
                           ifelse (pop == "c_ll_cr_n006","Carpathians", 
                           ifelse (pop == "c_ll_to-c_ll_ka-c_ll_og_n008", "Mongolia", 
                           ifelse (pop == "c_ll_tu_n006", "Tuva", 
                           ifelse (pop == "c_ll_vl_n008", "Vladivostok",  
                           ifelse (pop == "c_ll_ya_n008", "Yakutia", NA)))))))))))))))  %>% 
filter(Populations!="Balkans") 
  
diversity$watterson_ave <- as.numeric(gsub("\\*10\\^","e",diversity$watterson_ave))
diversity$pairwise_ave <- as.numeric(gsub("\\*10\\^","e",diversity$pairwise_ave))
diversity$tajimaD <- as.numeric(gsub("\\*10\\^","e",diversity$tajimaD))
  

diversity$Populations <- factor (diversity$Populations, levels=c("Norway", "Balkans", "Carpathians", "Bialowieza", "Latvia", "Kirov", "Urals", "Tuva", "Khentii", "Töv", "Ömnögovi", "Mongolia",  "Yakutia", "Vladivostok"))



diversity_to_plot<- data.frame(pop=character(),
                 Populations=character(), 
                 autosomic=double(),
                 X_chr=double(),
                 ratio=double(),
                 index=character(),
                 stringsAsFactors=FALSE) 

for(POP in unique(diversity$pop)){
diversity_watterson_per_pop <-  subset(diversity, pop==POP) %>% select (pop,region,watterson_ave,Populations) %>% spread(., region, watterson_ave) %>% mutate (ratio=X_chr/autosomic)  %>% mutate(index="watterson")
diversity_to_plot <- rbind(diversity_to_plot, diversity_watterson_per_pop)

diversity_pairwise_per_pop <-  subset(diversity, pop==POP) %>% select (pop,region,pairwise_ave,Populations) %>% spread(., region, pairwise_ave) %>% mutate (ratio=X_chr/autosomic)  %>% mutate(index="pairwise")
diversity_to_plot <- rbind(diversity_to_plot, diversity_pairwise_per_pop)

}
  
cols <- c("Carpathians"=brewer.pal(12,"Paired")[9], 
           "Kirov"=brewer.pal(12,"Paired")[1],  
           "Latvia"=brewer.pal(12,"Paired")[3],  
           "Norway"=brewer.pal(12,"Paired")[2],  
           "Bialowieza"=brewer.pal(12,"Paired")[4],  
           "Mongolia"=brewer.pal(12,"Paired")[12], 
           "Töv"=brewer.pal(12,"Paired")[12], 
          "Khentii"=brewer.pal(12,"Paired")[7], 
          "Ömnögovi"=brewer.pal(12,"Paired")[11], 
            "Tuva"=brewer.pal(12,"Paired")[8],  
           "Urals"=brewer.pal(11,"BrBG")[9],  
           "Vladivostok"=brewer.pal(12,"Paired")[5], 
           "Yakutia"=brewer.pal(12,"Paired")[6], 
           "Balkans"=brewer.pal(12,"Paired")[10])  

 
for (INDICE in unique(diversity_to_plot$index))
{
 ggplot (filter(diversity_to_plot,index==INDICE), aes(x=Populations, y=ratio, fill=Populations)) + 
   geom_point(color="dimgrey", size=5, shape=22) + 
   scale_fill_manual(values = cols) + 
  xlab ("Populations") + ylab (paste("Ratio X/autosome:",INDICE, sep="")) +  theme_classic() + 
theme (legend.text=element_text(size=18),legend.title = element_blank(), axis.title = element_text(size = 20), axis.text = element_text(size=18), title=element_text(size=25), axis.text.x = element_blank() ) 
ggsave(paste("/Users/marialucenaperez/Desktop/ratio_Xchr_autosomic_",INDICE,".pdf", sep=""), device = pdf,  width = 20, height = 10, units = "cm") 

 }






