---
title: "pipeline_contemporary_sfs"
output: html_document
---

# Generating SAF for intergenic data

## Defining variables:

I checked that all bam files were indexed --> OK
Ruta a bam files intergenic:

Global variables:

```{r, engine=bash, eval=FALSE}

RUTA=/home/mlucena/ANGSD_analysis 

cd $RUTA/intergenic_analysis/intergenic_sfs

#To launch one by one

POP="c_ll_ur"  # <--CHANGE POP HERE

#screen -S "$POP"_sfs
screen -S "$POP"_thetas

POP="c_ll_ur"  # <--CHANGE POP HERE

#script "$POP"_sfs.log
script "$POP"_thetas.log

POP="c_ll_ur"  # <--CHANGE POP HERE

THREADS=25                     # no. of computer cores used by bwa and samtools. 20 = OK, >20 = ask people first!
RUTA=/home/mlucena/ANGSD_analysis/depth_calculus 

read POP mean sd mean_truncated sd_truncated maxDepth minDepth maxDepth_truncated minDepth_truncated < $RUTA/"${POP}"_n*_mean_sd_depthGlobal_lynx_per_pop_mean_folds_0.95.csv


ANGSD="/opt/angsd/angsd"
NGSTOOLS="/opt/angsd/angsd/misc"
REF="/home/GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/lp23_without_repetitive_transposable_low_complexity.fa"
ANC="/home/GRUPOS/grupolince/reference_genomes/lynx_rufus_genome/c_lr_zz_0001_recal1.fa"
FILTER1=" -uniqueOnly 1 -remove_bads 1 -only_proper_pairs 1 -baq 1 -C 50 "
FILTER2=" -minMapQ 30 -minQ 20 -doCounts 1 "
N_IND=$(echo ${POP: -3} )
MIN_IND=$(expr $N_IND / 2)

# Sanity checks:

echo $POP
echo $N_IND
echo $MIN_IND
echo $minDepth


```

## Unfolded SAF 

```{r, engine=bash, eval=FALSE}

#The read command will read one line at a time from mean_sd_depthGlobal_lynx_per_pop_sd_folds2.csv
# put the first field (fields being blank separated, and where backslash can be used to escape the field or line separator) in the variable $discard, the second in $x, the third in $y and the rest of the fields in $discard.
#while read POP mean sd mean_truncated sd_truncated maxDepth minDepth maxDepth_truncated minDepth_truncated; 
#do

# Anyway I will do it one by one:

##########################
#  SAF (likelihood):     
##########################

echo "-------$POP----------SAF (likelihood)-----------------------------------------"

$ANGSD/angsd -P $THREADS -b /home/mlucena/ANGSD_analysis/intergenic_analysis/"$POP".intergenic.bamlist -ref $REF -anc $ANC \
-out "$POP".unfolded-lr \
$FILTER1 \
$FILTER2 \
-GL 1 -doSaf 1  \
-minInd  $MIN_IND -setMaxDepth $maxDepth -setMinDepth $minDepth

```

## SFS 

```{r, engine=bash, eval=FALSE}

##########################
#  SFS                   #(I dont require the -rf file as the saf already only contains the -rf sites
##########################
echo "-------$POP----------SFS------------------------------------------------------"

$NGSTOOLS/realSFS "$POP".unfolded-lr.saf.idx  -P $THREADS > "$POP".unfolded-lr.sfs

```

##  SAF (postprob) 

```{r, engine=bash, eval=FALSE}


echo "-------$POP----------SAF (postprob)-----------------------------------------"

$ANGSD/angsd -P $THREADS -b /home/mlucena/ANGSD_analysis/intergenic_analysis/"$POP".intergenic.bamlist -ref $REF -anc $ANC \
-out "$POP".unfolded-lr.postprob \
$FILTER1 \
$FILTER2 \
-GL 1 -doSaf 1  \
-minInd  $MIN_IND -setMaxDepth $maxDepth \
-doThetas 1 -pest "$POP".unfolded-lr.sfs 

```




##  Make bed theta        

```{r, engine=bash, eval=FALSE}

echo "-------$POP----------Make bed theta -----------------------------------------"

$NGSTOOLS/thetaStat print $POP.unfolded-lr.postprob.thetas.idx > $POP.printed.stats

# Transform the coordinates to use 0Based bedtools

awk 'NR>1 {print  $1" "$2-1" "$2}' $POP.printed.stats > $POP.0basedcoordinates.borrar # --> 1, 2, 3
# log transformation of watterson and pairwise:
awk 'NR>1 {print  "e("$3")"}' $POP.printed.stats | bc -l > "$POP".watterson.borrar # --> 4
awk 'NR>1 {print  "e("$4")"}' $POP.printed.stats | bc -l > "$POP".pairwise.borrar # --> 5
                                                                                  # --> PAIRWISE - WATTERSON
awk 'NR>1 {print  "e("$5")"}' $POP.printed.stats | bc -l > "$POP".thetaFL.borrar # --> 6
awk 'NR>1 {print  "e("$6")"}' $POP.printed.stats | bc -l > "$POP".thetaH.borrar # --> 7
awk 'NR>1 {print  "e("$7")"}' $POP.printed.stats | bc -l > "$POP".thetaL.borrar # --> 8

# Create a header for the "$POP".transformedThetas file

echo "scaffold position1 position2 watterson pairwise pairwise-watterson thetaFL thetaH thetaL" > "$POP".transformedThetas

paste $POP.0basedcoordinates.borrar "$POP".watterson.borrar "$POP".pairwise.borrar "$POP".thetaFL.borrar "$POP".thetaH.borrar "$POP".thetaL.borrar | \
awk -v OFS='\t'  '{print $1, $2, $3, $4, $5, $5-$4, $6, $7, $8}' >> "$POP".transformedThetas



# Hasta aquí!

rm "$POP".*.borrar


scp *.printed.stats /backup/mlucena/intermediate_files_ANGSD/intergenic_analysis/



```



```{bash}
nano heterozigosity_calculus.R
```

```{bash}
#in R
library (dplyr)
args <- commandArgs(TRUE)
print(args)
a<-scan(args[1])
heterozigosity <- a[2]/sum(a)
write (heterozigosity,paste0(args[1],"_heterozigosity.csv"))

```

```{bash}


for i in *sfs;
do
echo $i
Rscript heterozigosity_calculus.R $i 
done

rm heterozigosity_per_pop.csv

for i in *heterozigosity_borrar.csv
do
echo $i
heterozigosity=$(cat $i)
echo ${i/.unfolded-lr.sfs_heterozigosity_borrar.csv/} $heterozigosity >> heterozigosity_per_pop.csv
done

rm *_heterozigosity_borrar.csv




```




## SFS graphical representation


First we copy the SFS files to our local folder.

```{bash}
scp mlucena@genomics-b.ebd.csic.es:/home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs/*sfs /Users/marialucenaperez/Dropbox/phd/contemporary/ANGSD/intergenic_analysis/analysis_sfs

# I need to change name as R does not like "-lr.sfs"

cd /Users/marialucenaperez/Dropbox/phd/contemporary/ANGSD/intergenic_analysis/analysis_sfs


for i in *sfs
do 
#echo $i
echo $i "--->" ${i/-lr.sfs/_lr.sfs}
mv $i ${i/-lr.sfs/_lr.sfs}
rm $i
done

## OJO!!!!! READMEEEEEE!!!!! #######

# Le he quitado a mano a c_ll_to-c_ll_ka-c_ll_og_n008.unfolded_lr.sfs los guiones medios porque si no, no corre en R!!

```

Now we run the script. 

 
```{bash}
nano SFS_graphical_representation
```

```{r}
library(dplyr)
library(magrittr)

wd<-"/Users/marialucenaperez/Dropbox/phd/contemporary/ANGSD/intergenic_analysis/analysis_sfs/"

my_files_unfolded = list.files(path = wd, pattern="*sfs$")    
for (i in 1:length(my_files_unfolded)) 
{ assign(my_files_unfolded[i], (scan(paste(wd,my_files_unfolded[i], sep = ""), sep = " ", dec = ".")) %>% .[!is.na(.)])}

norm <- function(x) x/sum(x)

for (i in 1:length (my_files_unfolded))
{
  temp1=(eval(parse(text=my_files_unfolded[i])))
  temp2=norm(temp1[-1])
  temp3=norm(temp2[-c(length(temp2))])
  barplot(temp3,xlab="Alleles",
          names=1:length(temp3),ylab="Proportions",main=my_files_unfolded[i],col='grey')
  dev.print(device=postscript, paste("/Users/marialucenaperez/Dropbox/phd/contemporary/ANGSD/intergenic_analysis/analysis_sfs/",my_files_unfolded[i],".eps", sep=""), onefile=FALSE, horizontal=FALSE)
  dev.off()
  rm(temp1)
  rm(temp2)
  rm(temp3)
}

```

# SAF variable_sites 1

As we find that balkans have a wierd behaviour, we are going to repeat all the stats, using onel sites that were already filter out by SNP-pval 1e-4 for PCA and ADMIXTURE. We found that this threshold is already stable, and therefore, SNPs called are trustworthy. 

First we are going to select this sites, using PCA maf file.


```{bash}
# This is how it looks:

mkdir /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs/intergenic_sfs_filtered_sites

cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_PCA
zcat c_ll_no-c_ll_ba-h_ll_ba-c_ll_cr-c_ll_po-c_ll_la-c_ll_ki-c_ll_ur-c_ll_tu-c_ll_to-c_ll_ka-c_ll_og-c_ll_ya-c_ll_vl_n080.intergenic.mafs.gz | head
# chromo	position	major	minor	ref	knownEM	pK-EM	nInd
# lp23.s00001	3926	G	A	N	0.020146	9.075109e-08	80
# lp23.s00001	5895	T	A	N	0.031340	3.633791e-08	80
# lp23.s00001	6927	C	T	C	0.015890	0.000000e+00	80
# lp23.s00001	9059	C	T	N	0.064255	0.000000e+00	80
# lp23.s00001	13418	G	A	A	0.065885	0.000000e+00	80
# lp23.s00001	13869	T	G	T	0.025897	0.000000e+00	80
# lp23.s00001	14821	T	G	T	0.041811	0.000000e+00	80

# We are making a site file using the first column (See ANGSD documentation)

cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs/intergenic_sfs_filtered_sites

zcat /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_PCA/c_ll_no-c_ll_ba-h_ll_ba-c_ll_cr-c_ll_po-c_ll_la-c_ll_ki-c_ll_ur-c_ll_tu-c_ll_to-c_ll_ka-c_ll_og-c_ll_ya-c_ll_vl_n080.intergenic.mafs.gz | awk  'BEGIN{OFS="\t";} ; NR>1{print $1,$2}' > PCA_Maf1_SNP-pval1e_4.sites

angsd sites index PCA_Maf1_SNP-pval1e_4.sites

```


```{bash}

screen -S diversity_calculus_subset

POPS=$(ls /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs/*.transformedThetas | cut -d"/" -f 7)


echo -e "pop\tpairwise\twatterson"  > Maf1_SNP-pval1e_4_diversity_averages_per_pop.tsv

for POP in ${POPS[@]}

do

echo "grep for" $POP

grep -wFf PCA_Maf1_SNP-pval1e_4.sites /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs/$POP > ${POP/.transformedThetas/_Maf1_SNP-pval1e_4.transformedThetas.borrar}

WATTERSON_AVERAGE=$(cut -f 4 ${POP/.transformedThetas/_Maf1_SNP-pval1e_4.transformedThetas.borrar} | awk -v OFS='\t' '{for(i=1;i<=NF;i++) {sum[i] += $i }} END {for (i=1;i<=NF;i++) {printf ("%.10e\n", sum[i]/NR )}}' | sed 's/[eE]+\{0,1\}/*10^/g')

PAIRWISE_AVERAGE=$(cut -f 5 ${POP/.transformedThetas/_Maf1_SNP-pval1e_4.transformedThetas.borrar} | awk -v OFS='\t' '{for(i=1;i<=NF;i++) {sum[i] += $i }} END {for (i=1;i<=NF;i++) {printf ("%.10e\n", sum[i]/NR )}}' | sed 's/[eE]+\{0,1\}/*10^/g') 

DIFFERENCE_PAIRWISE_WATTERSON_SD=$(cut -f 6  ${POP/.transformedThetas/_Maf1_SNP-pval1e_4.transformedThetas.borrar} |  awk -v OFS='\t' '{for(i=1;i<=NF;i++) {sum[i] += $i; sumsq[i] += ($i)^2}} END {for (i=1;i<=NF;i++) {printf ("%.10e\n", sqrt((sumsq[i]-sum[i]^2/NR)/NR ))}}' | sed 's/[eE]+\{0,1\}/*10^/g') 

TAJIMAS_D=$(echo "(($PAIRWISE_AVERAGE) - ($WATTERSON_AVERAGE))/($DIFFERENCE_PAIRWISE_WATTERSON_SD)" | bc -l | awk '{printf ("%.10e\n",$1)}' |  sed 's/[eE]+\{0,1\}/*10^/g' )

paste \
<(echo ${POP/.transformedThetas/}) \
<(echo $WATTERSON_AVERAGE) \
<(echo $PAIRWISE_AVERAGE) \
<(echo $TAJIMAS_D) |\
sed 's/ /\t/g'| sed 's/\t\+/\t/g' >> Maf1_SNP-pval1e_4_diversity_averages_per_pop.tsv

rm *borrar

done





```


# SAF variable_sites 2

```{r, engine=bash, eval=FALSE}

cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_sfs_filtered_sites

#To launch one by one

POP="c_ll_ur"  # <--CHANGE POP HERE

#screen -S "$POP"_sfs
screen -S "$POP"_thetas_filtered

POP="c_ll_ur"  # <--CHANGE POP HERE

#script "$POP"_sfs.log
script "$POP"_thetas_filtered.log

POP="c_ll_ur"  # <--CHANGE POP HERE

THREADS=25                     # no. of computer cores used by bwa and samtools. 20 = OK, >20 = ask people first!
RUTA=/home/mlucena/ANGSD_analysis/depth_calculus 

read POP mean sd mean_truncated sd_truncated maxDepth minDepth maxDepth_truncated minDepth_truncated < $RUTA/"${POP}"_n*_mean_sd_depthGlobal_lynx_per_pop_mean_folds_0.95.csv

ANGSD="/opt/angsd/angsd"
NGSTOOLS="/opt/angsd/angsd/misc"
REF="/home/GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/lp23_without_repetitive_transposable_low_complexity.fa"
ANC="/home/GRUPOS/grupolince/reference_genomes/lynx_rufus_genome/c_lr_zz_0001_recal1.fa"
FILTER1=" -uniqueOnly 1 -remove_bads 1 -only_proper_pairs 1 -baq 1 -C 50 "
FILTER2=" -minMapQ 30 -minQ 20 -doCounts 1 "
N_IND=$(echo ${POP: -3} )
MIN_IND=$(expr $N_IND / 2)
SITES="PCA_Maf1_SNP-pval1e_4.sites"


# Sanity checks:

echo $POP
echo $N_IND
echo $MIN_IND
echo $minDepth
echo $SITES

```

## Unfolded SAF 

```{r, engine=bash, eval=FALSE}

#The read command will read one line at a time from mean_sd_depthGlobal_lynx_per_pop_sd_folds2.csv
# put the first field (fields being blank separated, and where backslash can be used to escape the field or line separator) in the variable $discard, the second in $x, the third in $y and the rest of the fields in $discard.
#while read POP mean sd mean_truncated sd_truncated maxDepth minDepth maxDepth_truncated minDepth_truncated; 
#do

# Anyway I will do it one by one:

##########################
#  SAF (likelihood):     
##########################

echo "-------$POP----------SAF (likelihood)-----------------------------------------"

$ANGSD/angsd -P $THREADS -b /home/mlucena/ANGSD_analysis/intergenic_analysis/"$POP".intergenic.bamlist -ref $REF -anc $ANC \
-out "$POP"_filtered_Maf1_SNP-pval1e_4.unfolded_lr \
$FILTER1 \
$FILTER2 \
-sites $SITES \
-GL 1 -doSaf 1  \
-minInd  $MIN_IND -setMaxDepth $maxDepth -setMinDepth $minDepth

```

Sanity check

```{bash}

$NGSTOOLS/realSFS print  c_ll_ur_n006_filtered_Maf1_SNP-pval1e_4.unfolded_lr.saf.idx | wc -l
# 2058550

wc -l PCA_Maf1_SNP-pval1e_4.sites
# 2096552 PCA_Maf1_SNP-pval1e_4.sites

$NGSTOOLS/realSFS print  c_ll_ur_n006_filtered_Maf1_SNP-pval1e_4.unfolded_lr.saf.idx | awk  'BEGIN{OFS="\t";} ;{print $1, $2}'  > sitios_SAF.sites

 wc -l sitios_SAF.sites 
# 2058550 sitios_SAF.sites


diff sitios_SAF.sites PCA_Maf1_SNP-pval1e_4.sites | head
# 23a24
# > lp23.s00001	53180
# 50a52
# > lp23.s00001	95912
# 57a60
# > lp23.s00001	120814
# 59a63,65
# > lp23.s00001	121922
# > lp23.s00001	121934
# > lp23.s00001	121970

# ¿Todos los sitios de sitios_SAF.sites están en PCA_Maf1_SNP-pval1e_4.sites?

diff sitios_SAF.sites PCA_Maf1_SNP-pval1e_4.sites | grep "<"  
#Nada! Ergo, sí. 


# Voy a buscar este sitio en el SAF de todo el genoma:
# lp23.s00001	53180

cd ../intergenic_sfs

$NGSTOOLS/realSFS print c_ll_ur_n006.unfolded-lr.saf.idx | grep "lp23.s00001" | grep "53180"


```



## SFS 

```{r, engine=bash, eval=FALSE}

##########################
#  SFS                   #(I dont require the -rf file as the saf already only contains the -rf sites
##########################
echo "-------$POP----------SFS------------------------------------------------------"

$NGSTOOLS/realSFS "$POP"_filtered_Maf1_SNP-pval1e_4.unfolded_lr.saf.idx  -P $THREADS > "$POP"_filtered_Maf1_SNP-pval1e_4.unfolded_lr.sfs

```

##  SAF (postprob) 

```{r, engine=bash, eval=FALSE}


echo "-------$POP----------SAF (postprob)-----------------------------------------"

$ANGSD/angsd -P $THREADS -b /home/mlucena/ANGSD_analysis/intergenic_analysis/"$POP".intergenic.bamlist -ref $REF -anc $ANC \
-out "$POP"_filtered_Maf1_SNP-pval1e_4.unfolded_lr.postprob \
$FILTER1 \
$FILTER2 \
-GL 1 -doSaf 1  \
-sites $SITES \
-minInd  $MIN_IND -setMaxDepth $maxDepth \
-doThetas 1 -pest "$POP"_filtered_Maf1_SNP-pval1e_4.unfolded_lr.sfs 

```




##  Make bed theta        

```{r, engine=bash, eval=FALSE}

echo "-------$POP----------Make bed theta -----------------------------------------"

$NGSTOOLS/thetaStat print "$POP"_filtered_Maf1_SNP-pval1e_4.unfolded_lr.postprob.thetas.idx > "$POP"_filtered_Maf1_SNP-pval1e_4.printed.stats

# Transform the coordinates to use 0Based bedtools

awk 'NR>1 {print  $1" "$2-1" "$2}' "$POP"_filtered_Maf1_SNP-pval1e_4.printed.stats > "$POP"_filtered_Maf1_SNP-pval1e_4.0basedcoordinates.borrar # --> 1, 2, 3
# log transformation of watterson and pairwise:
awk 'NR>1 {print  "e("$3")"}' "$POP"_filtered_Maf1_SNP-pval1e_4.printed.stats | bc -l > "$POP"_filtered_Maf1_SNP-pval1e_4.watterson.borrar # --> 4
awk 'NR>1 {print  "e("$4")"}' "$POP"_filtered_Maf1_SNP-pval1e_4.printed.stats | bc -l > "$POP"_filtered_Maf1_SNP-pval1e_4.pairwise.borrar # --> 5
                                                                                  # --> PAIRWISE - WATTERSON
awk 'NR>1 {print  "e("$5")"}' "$POP"_filtered_Maf1_SNP-pval1e_4.printed.stats | bc -l > "$POP"_filtered_Maf1_SNP-pval1e_4.thetaFL.borrar # --> 6
awk 'NR>1 {print  "e("$6")"}' "$POP"_filtered_Maf1_SNP-pval1e_4.printed.stats | bc -l > "$POP"_filtered_Maf1_SNP-pval1e_4.thetaH.borrar # --> 7
awk 'NR>1 {print  "e("$7")"}' "$POP"_filtered_Maf1_SNP-pval1e_4.printed.stats | bc -l > "$POP"_filtered_Maf1_SNP-pval1e_4.thetaL.borrar # --> 8

# Create a header for the "$POP"_filtered_Maf1_SNP-pval1e_4.transformedThetas file

echo "scaffold position1 position2 watterson pairwise pairwise-watterson thetaFL thetaH thetaL" > "$POP"_filtered_Maf1_SNP-pval1e_4.transformedThetas

paste "$POP"_filtered_Maf1_SNP-pval1e_4.0basedcoordinates.borrar "$POP"_filtered_Maf1_SNP-pval1e_4.watterson.borrar "$POP"_filtered_Maf1_SNP-pval1e_4.pairwise.borrar "$POP"_filtered_Maf1_SNP-pval1e_4.thetaFL.borrar "$POP"_filtered_Maf1_SNP-pval1e_4.thetaH.borrar "$POP"_filtered_Maf1_SNP-pval1e_4.thetaL.borrar | \
awk -v OFS='\t'  '{print $1, $2, $3, $4, $5, $5-$4, $6, $7, $8}' >> "$POP"_filtered_Maf1_SNP-pval1e_4.transformedThetas



# Hasta aquí!

rm "$POP"_filtered_Maf1_SNP-pval1e_4.*.borrar


scp *.printed.stats /backup/mlucena/intermediate_files_ANGSD/intergenic_analysis/


  
  



```



