---
title: "4.Windows_per_feature"
output: html_document
---

Test to see whether we could concatenate all the features and run a window approach over it.


# 1. Get position for every feature:

```{bash}

mkdir /home/mlucena/ANGSD_analysis/whole_genome_analysis/sfs/diversity_window_feature

cd /home/mlucena/ANGSD_analysis/whole_genome_analysis/sfs/

POP=c_lp_do_n012 # <--- Change pop here!
screen -S "$POP"_get_feature_coordinates
POP=c_lp_do_n012 # <--- Change pop here!
script log_screen_"$POP"_get_feature_coordinates.log
POP=c_lp_do_n012 # <--- Change pop here!

cd /home/mlucena/ANGSD_analysis/whole_genome_analysis/sfs

## Variables
SCAFFOLDS_FOLDER=/GRUPOS/grupolince/Lyp_annotation_Apr14_final/LYPA23C.CDS.GENE_promoters.GENE_introns.UTRs.ncRNA.lncRNA.lncRNA_introns.lncRNA_promoters.UCNE.intergenic.nr.gff3.PerScaffold/
FEATURE_POSITIONS_FOLDER=/home/mlucena/ANGSD_analysis/whole_genome_analysis/sfs/diversity_window_feature/
SFS_FOLDER=/home/mlucena/ANGSD_analysis/whole_genome_analysis/sfs/
FEATURE="CDS"
OUTPUT_DIR=/home/mlucena/ANGSD_analysis/whole_genome_analysis/sfs/diversity_window_feature
REF="/GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated"

FEATURES=(UCNE intergenic 3UTR 5UTR intron intron_lncRNA lncRNA ncRNA promoter_gene_1000 promoter_lncRNA_1000)

echo "---------------------------------------------------$POP---------------------------------------------------"

for FEATURE in "${FEATURES[@]}"
do
echo $FEATURE
for iteration in {1..10}
do
rm "${OUTPUT_DIR}"/"${POP}".transformedThetas_iteration"$iteration"."${FEATURE}"
SCAFFOLD_START=$( echo '1 + 4170 * ('$iteration' - 1)' | bc)
SCAFFOLD_END=$( echo '4170 * '$iteration | bc )
for SCAFFOLD_NUMBER in $(seq -s " " -f %05g $SCAFFOLD_START $SCAFFOLD_END) # esta manera de hacer seq te va a sacar siempre 5 cifras.
do
SCAFFOLD=$(echo "lp23.s"$SCAFFOLD_NUMBER)
cd "$SFS_FOLDER""$POP"_separated_by_scaffold
cat "$POP".transformedThetas_"$SCAFFOLD" | bedtools intersect -a stdin -b <( cat "$SCAFFOLDS_FOLDER"LYPA23C.CDS.GENE_promoters.GENE_introns.UTRs.ncRNA.lncRNA.lncRNA_introns.lncRNA_promoters.UCNE.intergenic.nr.gff3_"$SCAFFOLD" | grep $FEATURE) -sorted -g $REF >> "${OUTPUT_DIR}"/"${POP}".transformedThetas_iteration"$iteration"."${FEATURE}"
done &
done 
done

```

# 2. Get common positions among pops

We are using the intermediate interaction files to avoid large files.

## Bedtools approach

```{bash}

cd /home/mlucena/ANGSD_analysis/whole_genome_analysis/sfs/diversity_window_feature

REF="/GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated"
OUTPUT_DIR=/home/mlucena/ANGSD_analysis/whole_genome_analysis/sfs/diversity_window_feature
FEATURES=(5UTR intron intron_lncRNA lncRNA ncRNA promoter_gene_1000 promoter_lncRNA_1000)

#3UTR intergenic UCNE

for FEATURE in "${FEATURES[@]}"
do
echo $FEATURE
for iteration in {1..10}
do
echo $iteration
echo "doing lynx pardinus"
bedtools intersect -a <(cat c_lp_sm_n012.transformedThetas_iteration"$iteration"."${FEATURE}" |  awk -v OFS='\t' '{print $1, $2, $3}') -b <(cat c_lp_do_n012.transformedThetas_iteration"$iteration"."${FEATURE}" |  awk -v OFS='\t' '{print $1, $2, $3}') -sorted -g $REF > c_lp_iteration"$iteration"."${FEATURE}".innercoordinates

# Filter do for the coordinates in common.
bedtools intersect -a  <(cat "${OUTPUT_DIR}"/c_lp_do_n012.transformedThetas_iteration"$iteration"."${FEATURE}") -b <(cat c_lp_iteration"$iteration"."${FEATURE}".innercoordinates) -sorted -g $REF > "${OUTPUT_DIR}"/c_lp_do_n012.transformedThetas_iteration"$iteration".filtered."${FEATURE}"

# Filter sm for the coordinates in common.
bedtools intersect -a  <(cat "${OUTPUT_DIR}"/c_lp_do_n012.transformedThetas_iteration"$iteration"."${FEATURE}") -b <(cat c_lp_iteration"$iteration"."${FEATURE}".innercoordinates) -sorted -g $REF> "${OUTPUT_DIR}"/c_lp_sm_n012.transformedThetas_iteration"$iteration".filtered."${FEATURE}"

echo "doing lynx lynx"

## LYNX LYNX ## ## ## ## ## 
# I am doing a inner join:
# I am doing it in 3 different steps althougth I would need it, because, when I get to ll I will need to do it like that, or it will be more complex.
# 1. Get coordinates in common for the three pop.
bedtools intersect -a <(cat c_ll_ki_n008.transformedThetas_iteration"$iteration"."${FEATURE}" |  awk -v OFS='\t' '{print $1, $2, $3}') -b <(cat c_ll_po_n008.transformedThetas_iteration"$iteration"."${FEATURE}" |  awk -v OFS='\t' '{print $1, $2, $3}') -sorted -g $REF | bedtools intersect -a - -b <(cat c_ll_no_n008.transformedThetas_iteration"$iteration"."${FEATURE}" |  awk -v OFS='\t' '{print $1, $2, $3}') -sorted -g $REF> c_ll_iteration"$iteration"."${FEATURE}".innercoordinates

# Filter ki for the coordinates in common.
bedtools intersect -a  <(cat "${OUTPUT_DIR}"/c_ll_ki_n008.transformedThetas_iteration"$iteration"."${FEATURE}") -b <(cat c_ll_iteration"$iteration"."${FEATURE}".innercoordinates) -sorted -g $REF > "${OUTPUT_DIR}"/c_ll_ki_n008.transformedThetas_iteration"$iteration".filtered."${FEATURE}"

# Filter no for the coordinates in common.
bedtools intersect -a  <(cat "${OUTPUT_DIR}"/c_ll_no_n008.transformedThetas_iteration"$iteration"."${FEATURE}") -b <(cat c_ll_iteration"$iteration"."${FEATURE}".innercoordinates) -sorted -g $REF > "${OUTPUT_DIR}"/c_ll_no_n008.transformedThetas_iteration"$iteration".filtered."${FEATURE}"

# Filter po for the coordinates in common.
bedtools intersect -a  <(cat "${OUTPUT_DIR}"/c_ll_po_n008.transformedThetas_iteration"$iteration"."${FEATURE}") -b <(cat c_ll_iteration"$iteration"."${FEATURE}".innercoordinates) -sorted -g $REF > "${OUTPUT_DIR}"/c_ll_po_n008.transformedThetas_iteration"$iteration".filtered."${FEATURE}"
done
done

wc -l c_lp*iteration*
wc -l c_ll*iteration*

# Para pardinus tiene buena pinta!
rm c_lp_iteration*.innercoordinates



FEATURES=(UCNE 3UTR intergenic 5UTR intron intron_lncRNA lncRNA ncRNA promoter_gene_1000 promoter_lncRNA_1000)

for FEATURE in "${FEATURES[@]}"
do
echo $FEATURE
rm c_ll_ki_n008.transformedThetas_iteration{1..10}."${FEATURE}"
rm c_ll_po_n008.transformedThetas_iteration{1..10}."${FEATURE}"
rm c_ll_no_n008.transformedThetas_iteration{1..10}."${FEATURE}"
rm c_lp_sm_n012.transformedThetas_iteration{1..10}."${FEATURE}"
rm c_lp_do_n012.transformedThetas_iteration{1..10}."${FEATURE}"
done


```

## X and autosomes file

I have to create a filter to filter for with the scaffolds associated to X chr and the scaffold associated with autosomes. 
```{bash}
cd /GRUPOS/grupolince/copia_fabascal/MAPPINGS/

awk '{print $1}' lynx2cat_wTiger_Xchr.sorted.merged.bed | sort | uniq > scaffold_Xchr_putatively
wc -l scaffold_Xchr_putatively
#1891 scaffold_X
# Those would be the scaffolds associated to Xchr, however! some of those are also present in autosomes, so I should filter them. 
grep -f scaffold_Xchr_putatively lynx2cat_wTiger_AUTOSOMALchr.sorted.merged.corrected.bed | awk '{print $1}' | sort | uniq > scaffold_X_Autosomes
wc -l scaffold_X_Autosomes
#250 scaffold_X_Autosomes

grep -v -f scaffold_X_Autosomes scaffold_Xchr > scaffold_Xchr_exclusively
wc -l scaffold_Xchr_exclusively
#1641 scaffold_Xchr_exclusively

# He guardado un README con esta info en la capeta. 

rm scaffold_X_Autosomes
# Vamos a usar el scaffold_Xchr_putatively para quitarnos todo lo que pudiera ser X del analisis de autosomico y scaffold_Xchr_exclusively para hacer el análisis de los X. 

# As I will use them with bedtools intersect, I need to add some coordinates. To do so I will grep those scaffold in lp23.lenght file and add the complete chr.

grep -f /GRUPOS/grupolince/copia_fabascal/MAPPINGS/scaffold_Xchr_exclusively /GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated > /GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated_Xchr


grep -v -f /GRUPOS/grupolince/copia_fabascal/MAPPINGS/scaffold_Xchr_putatively /GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated > /GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated_autosomes


```

As I thought about this filter after doing already some features, I have to different approaches right now. 

1. Filter the scaffolds before cat (intron_lncRNA lncRNA ncRNA promoter_gene_1000 promoter_lncRNA_1000 3UTR UCNE 5UTR intergenic CDS).
2. Filter the scaffolds already in the transformed filtered file (3UTR UCNE 5UTR intergenic CDS intron).

What I am doing is:

Using the scaffold only assing to X chr to get X coordinates. However! I will use scaffold assinged to X OR X&Autosomes to filter all the scaffold. 
i.e.

X= scaffolds Assing only to X 
Autosomes= All scaffolds - (scaffold assigned to x OR to x & autosomes).

Los archivos son:

/GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated_Xchr
/GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated_autosomes

# Join all filtered iteractions and assign consecutive coordinates to a file

```{bash}

cd /home/mlucena/ANGSD_analysis/whole_genome_analysis/sfs/diversity_window_feature

POPS=(c_ll_ki_n008 c_ll_po_n008 c_ll_no_n008 c_lp_do_n012 c_lp_sm_n012)

FEATURES=(intron_lncRNA lncRNA ncRNA promoter_gene_1000 promoter_lncRNA_1000 )
REF=/GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated


for FEATURE in "${FEATURES[@]}"
do
for POP in "${POPS[@]}"
do
# join all files
for iteration in {1..10}
do

echo "Doing "$FEATURE $POP $iteration 

bedtools intersect -a "$POP".transformedThetas_iteration$iteration.filtered."${FEATURE}" -b <(awk -v OFS="\t" '{print $1, "1", $2}' /GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated_Xchr) -sorted -g $REF > "$POP".transformedThetas_iteration$iteration.filtered.Xchr."${FEATURE}"

bedtools intersect -a "$POP".transformedThetas_iteration$iteration.filtered."${FEATURE}" -b <(awk -v OFS="\t" '{print $1, "1", $2}' /GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated_autosomes) -sorted -g $REF > "$POP".transformedThetas_iteration$iteration.filtered.Auto."${FEATURE}"

done
echo "finish filtering, ready to cat"
rm "$POP".transformedThetas_iteration{1..10}.filtered."${FEATURE}"

cat "$POP".transformedThetas_iteration{1..10}.filtered.Xchr."${FEATURE}" >> "$POP".transformedThetas.filtered.Xchr."${FEATURE}"
cat "$POP".transformedThetas_iteration{1..10}.filtered.Auto."${FEATURE}" >> "$POP".transformedThetas.filtered.Auto."${FEATURE}"

rm "$POP".transformedThetas_iteration{1..10}.filtered.Xchr."${FEATURE}"
rm "$POP".transformedThetas_iteration{1..10}.filtered.Auto."${FEATURE}"

echo "finish cat, ready to CONCAT"

# assign consecutive coordinates
awk -v OFS='\t' '{print "CONCATENATION", NR, NR+1, $4, $5, $6, $7, $8, $9}' "$POP".transformedThetas.filtered.Xchr."${FEATURE}" > "$POP".transformedThetas.filtered.CONCATENATION.Xchr."${FEATURE}" 

awk -v OFS='\t' '{print "CONCATENATION", NR, NR+1, $4, $5, $6, $7, $8, $9}' "$POP".transformedThetas.filtered.Auto."${FEATURE}" > "$POP".transformedThetas.filtered.CONCATENATION.Auto."${FEATURE}" 

echo "Done with "$POP" and" $FEATURE"

done
done

wc -l *CONCATENATION*
```

# Filtering the already done FEATURES


```{bash}
cd /home/mlucena/ANGSD_analysis/whole_genome_analysis/sfs/diversity_window_feature

POPS=(c_ll_ki_n008 c_ll_po_n008 c_ll_no_n008 c_lp_do_n012 c_lp_sm_n012)

FEATURES=(3UTR UCNE 5UTR intergenic CDS intron) 

REF=/GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated


for FEATURE in "${FEATURES[@]}"
do
for POP in "${POPS[@]}"
do
echo $POP $FEATURE "doing intersect"
echo "Xchr"
bedtools intersect -a "$POP".transformedThetas.filtered."${FEATURE}" -b <(awk -v OFS="\t" '{print $1, "1", $2}' /GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated_Xchr) -sorted -g $REF > "$POP".transformedThetas.filtered.Xchr."${FEATURE}"
echo "Autosomes"
bedtools intersect -a "$POP".transformedThetas.filtered."${FEATURE}" -b <(awk -v OFS="\t" '{print $1, "1", $2}' /GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/Length_scaffolds_lp23_tab_separated_autosomes) -sorted -g $REF > "$POP".transformedThetas.filtered.Auto."${FEATURE}"

echo $POP $FEATURE "doing concat"
echo "Xchr"
# assign consecutive coordinates
awk -v OFS='\t' '{print "CONCATENATION", NR, NR+1, $4, $5, $6, $7, $8, $9}' "$POP".transformedThetas.filtered.Xchr."${FEATURE}" > "$POP".transformedThetas.filtered.CONCATENATION.Xchr."${FEATURE}" 
echo "Autosomes"
awk -v OFS='\t' '{print "CONCATENATION", NR, NR+1, $4, $5, $6, $7, $8, $9}' "$POP".transformedThetas.filtered.Auto."${FEATURE}" > "$POP".transformedThetas.filtered.CONCATENATION.Auto."${FEATURE}" 

done
done

# BORRAR!
rm filtered."${FEATURE}"

```

# HASTA AQUI TENGO LANZADO!

# Calculate diversity per windows.

Over the files I will performe window summary using this awk script I found online (and tested that it worked). 
window=50000
slide=10000

```{bash}

POPS=(c_ll_ki_n008 c_ll_po_n008 c_ll_no_n008 c_lp_do_n012 c_lp_sm_n012)
FEATURES=(intergenic CDS intron intron_lncRNA lncRNA ncRNA promoter_gene_1000 promoter_lncRNA_1000 UCNE 3UTR 5UTR)

for FEATURE in "${FEATURES[@]}"
do
for POP in "${POPS[@]}"
do 

echo $POP $FEATURE "autosome"
awk -v OFS="\t" 'BEGIN{window=10000;slide=5000} { if(NR==1) print "coord_end","ave_watterson","ave_pairwise"} {mod=NR%window; if(NR<=window){count++}else{sum-=array[mod];sum2-=array2[mod]}sum+=$4;sum2+=$5;array[mod]=$4;array2[mod]=$5;} (NR%slide)==0{print NR,sum/count,sum2/count}' <(cat $POP.transformedThetas.filtered.CONCATENATION.Auto.$FEATURE) > $POP.transformedThetas.filtered.windows.AUTO.$FEATURE

echo $POP $FEATURE  "Xchr"
awk -v OFS="\t" 'BEGIN{window=10000;slide=5000} { if(NR==1) print "coord_end","ave_watterson","ave_pairwise"} {mod=NR%window; if(NR<=window){count++}else{sum-=array[mod];sum2-=array2[mod]}sum+=$4;sum2+=$5;array[mod]=$4;array2[mod]=$5;} (NR%slide)==0{print NR,sum/count,sum2/count}' <(cat $POP.transformedThetas.filtered.CONCATENATION.Xchr.$FEATURE) > $POP.transformedThetas.filtered.windows.XCHR.$FEATURE

done &
done

```

# Join per pop

```{bash}

POPS=(c_ll_ki_n008 c_ll_po_n008 c_ll_no_n008 c_lp_do_n012 c_lp_sm_n012)

for POP in "${POPS[@]}"
do 
rm $POP.transformedThetas.filtered.windows.AUTO
rm $POP.transformedThetas.filtered.windows.XCHR

echo $POP
for FEATURE in "${FEATURES[@]}"
do
echo $FEATURE
awk -v OFS="\t" -v FEATURE="$FEATURE" -v POP="$POP" ' { if(NR==1) print "coord_end","ave_watterson","ave_pairwise", "feature", "pop"} {print $0, FEATURE, POP}' <(tail -n+2 $POP.transformedThetas.filtered.windows.Auto.$FEATURE) >> $POP.transformedThetas.filtered.windows.AUTO
awk -v OFS="\t" -v FEATURE="$FEATURE" -v POP="$POP" ' { if(NR==1) print "coord_end","ave_watterson","ave_pairwise", "feature", "pop"} {print $0, FEATURE, POP}' <(tail -n+2 $POP.transformedThetas.filtered.windows.XCHR.$FEATURE) >> $POP.transformedThetas.filtered.windows.XCHR

done
done
```


# Download to local

```{bash}
scp mlucena@genomics-b.ebd.csic.es:/home/mlucena/ANGSD_analysis/whole_genome_analysis/sfs/diversity_window_feature/*.transformedThetas.filtered.windows /Users/marialucenaperez/Documents/WG_lynx_diversity_per_windows_features/raw
```

# R
# Create dataframes

## Load library & wd
```{r}
library(viridis)
library(ggplot2); theme_set(theme_minimal())
library(tidyr)
library(plyr)
library(dplyr)
library(data.table)
library(arules)
library(psych)
library(ggpubr)
library(boot)
library(broom)
library(ggpubr)
library(ggrepel)
library(ggpmisc)
library(gridExtra)
library(mgcv)
library(MuMIn)

wd_output <- "/Users/marialucenaperez/Documents/WG_lynx_diversity_per_windows_features/"
wd_input <- "/Users/marialucenaperez/Documents/WG_lynx_diversity_per_windows_features/raw/"
```

## Load pre-data

```{r}

c_ll_ki_n008_diversity_windows <- read.csv(paste(wd_input, "c_ll_ki_n008.transformedThetas.filtered.windows.UCNE", sep=""), header=T, na.strings=c("NA", "na"), stringsAsFactors=F, row.names=NULL, sep="\t", dec=".") %>% mutate (., color =  viridis_pal()(5)[1]) %>% mutate (feature="UCNE") %>% mutate (pop="c_ll_ki_n008") %>% mutate (Population="Kirov")

c_ll_no_n008_diversity_windows <- read.csv(paste(wd_input, "c_ll_no_n008.transformedThetas.filtered.windows.UCNE", sep=""), header=T, na.strings=c("NA", "na"), stringsAsFactors=F, row.names=NULL, sep="\t", dec=".") %>% mutate (., color = viridis_pal()(5)[2]) %>% mutate (feature="UCNE") %>% mutate (pop="c_ll_no_n008") %>% mutate (Population="Norway")

c_ll_po_n008_diversity_windows <- read.csv(paste(wd_input, "c_ll_po_n008.transformedThetas.filtered.windows.UCNE", sep=""), header=T, na.strings=c("NA", "na"), stringsAsFactors=F, row.names=NULL, sep="\t", dec=".") %>% mutate (., color =  viridis_pal()(5)[3]) %>% mutate (feature="UCNE") %>% mutate (pop="c_ll_po_n008") %>% mutate (Population="Poland")

c_lp_sm_n012_diversity_windows <- read.csv(paste(wd_input, "c_lp_sm_n012.transformedThetas.filtered.windows.UCNE", sep=""), header=T, na.strings=c("NA", "na"), stringsAsFactors=F, row.names=NULL, sep="\t", dec=".") %>% mutate (., color =  "#5DC863FF") %>% mutate (feature="UCNE") %>% mutate (pop="c_lp_sm_n008") %>% mutate (Population="Andujar")

c_lp_do_n012_diversity_windows <- read.csv(paste(wd_input, "c_lp_do_n012.transformedThetas.filtered.windows.UCNE", sep=""), header=T, na.strings=c("NA", "na"), stringsAsFactors=F, row.names=NULL, sep="\t", dec=".") %>% mutate (., color =  "#FDE725FF") %>% mutate (feature="UCNE") %>% mutate (pop="c_lp_do_n008") %>% mutate (Population="Donana")

```

# Create comparison dataframe.

```{r}

create_data_diversity_bootleneck_vs_non_bootleneck <- function(POP1, POP2, name_POP1, name_POP2){
    POPULATION1=POP1 %>% select(pop) %>% mutate(Populations=as.character(pop)) %>% .[1,1] 
    POPULATION2=POP2 %>% select(pop) %>% mutate(Populations=as.character(pop)) %>% .[1,1] 
  data_diversity_POP1_POP2 <- dplyr::inner_join (POP1, POP2, by = c("coord_end", "feature")) %>%  
      # Calculo el delta 
    dplyr::mutate (delta_pairwise = (ave_pairwise.y - ave_pairwise.x)) %>%
    dplyr::mutate (delta_watterson = (ave_watterson.y - ave_watterson.x)) %>% 
    dplyr::mutate (delta_pairwise_corrected = (ave_pairwise.y - ave_pairwise.x)/(ave_pairwise.y + ave_pairwise.x) ) %>% 
    dplyr::mutate (delta_watterson_corrected = ((ave_watterson.y - ave_watterson.x)/(ave_watterson.y + ave_watterson.x))) %>% 
       # Add comparison title  
    dplyr::mutate(comparison=paste(POPULATION1, "-", POPULATION2, sep=""))
  dataframename <- paste (name_POP1, name_POP2, sep="_")
  assign (dataframename, data_diversity_POP1_POP2,.GlobalEnv)
}

# Create data frame
# Dataframe Kirov-Norway
create_data_diversity_bootleneck_vs_non_bootleneck(c_ll_ki_n008_diversity_windows,c_ll_no_n008_diversity_windows, deparse(substitute(c_ll_ki_n008_diversity_windows)),deparse(substitute(c_ll_no_n008_diversity_windows)))
# Dataframe Kirov-Poland
create_data_diversity_bootleneck_vs_non_bootleneck(c_ll_ki_n008_diversity_windows,c_ll_po_n008_diversity_windows, deparse(substitute(c_ll_ki_n008_diversity_windows)),deparse(substitute(c_ll_po_n008_diversity_windows)))
# Dataframe Sierra_Morena-Doñana
create_data_diversity_bootleneck_vs_non_bootleneck(c_lp_sm_n012_diversity_windows,c_lp_do_n012_diversity_windows, deparse(substitute(c_lp_sm_n012_diversity_windows)),deparse(substitute(c_lp_do_n012_diversity_windows)))

# Combine comparisons
all_comparisons_diversity_bootleneck_vs_non_bootleneck_window <- rbind (c_ll_ki_n008_diversity_windows_c_ll_no_n008_diversity_windows, c_ll_ki_n008_diversity_windows_c_ll_po_n008_diversity_windows, c_lp_sm_n012_diversity_windows_c_lp_do_n012_diversity_windows)

```

## *Delta distribution
```{r}
DATAFRAMES=c("c_ll_ki_n008_diversity_windows_c_ll_no_n008_diversity_windows", "c_ll_ki_n008_diversity_windows_c_ll_po_n008_diversity_windows", "c_lp_sm_n019_diversity_windows_c_lp_do_n012_diversity_windows")

for (DATAFRAME in DATAFRAMES) 
{
  POP1=get(DATAFRAME) %>% select(pop.x) %>% mutate(pop.x=as.character(pop.x)) %>% .[1,1] 
  POP2=get(DATAFRAME) %>% select(pop.y) %>% mutate(pop.y=as.character(pop.y)) %>% .[1,1] 
  POPULATION1=get(DATAFRAME) %>% select(Population.x) %>% mutate(Population.x=as.character(Population.x)) %>% .[1,1] 
  POPULATION2=get(DATAFRAME) %>% select(Population.y) %>% mutate(Population.y=as.character(Population.y)) %>% .[1,1] 
  
 # Watterson
 # Plot de la distribución de delta
  ggplot(get(DATAFRAME), aes(x=(delta_watterson))) + 
    geom_histogram() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggsave (paste(wd_output,DATAFRAME,"_delta_watterson_average_distribution.pdf", sep=""))
  # Plot de la distribución de delta para cada feature
 ggplot() + 
    aes(x = delta_watterson) +
    geom_histogram() +
    facet_wrap(~feature, scales = "free")  +
    theme(axis.text.x = element_text(angle=90, hjust=1)) +
    ggsave (paste(wd_output,DATAFRAME,"_delta_watterson_average_distribution_per_feature.pdf", sep=""), width = 25, height = 25, units = "cm")
 
  # Plot de la distribución de delta corregida
  ggplot(get(DATAFRAME), aes(x=(delta_watterson_corrected))) + 
    geom_histogram() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggsave (paste(wd_output,DATAFRAME,"_delta_watterson_corrected_average_distribution.pdf", sep=""))
  # Plot de la distribución de delta para cada feature
 ggplot() + 
    aes(x = delta_watterson_corrected) +
    geom_histogram() +
    facet_wrap(~feature, scales = "free")  +
    theme(axis.text.x = element_text(angle=90, hjust=1)) +
    ggsave (paste(wd_output,DATAFRAME,"_delta_watterson_corrected_average_distribution_per_feature.pdf", sep=""), width = 25, height = 25, units = "cm")

 
  # PI
   # Distribution 
  ggplot(get(DATAFRAME), aes(x=(delta_pairwise))) + 
    geom_histogram() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggsave (paste(wd_output,DATAFRAME,"_delta_pairwise_average_distribution.pdf", sep=""))
  # Plot de la distribución de delta para cada feature
 ggplot() + 
    aes(x = delta_pairwise) +
    geom_histogram() +
    facet_wrap(~feature, scales = "free")  +
    theme(axis.text.x = element_text(angle=90, hjust=1)) +
    ggsave (paste(wd_output,DATAFRAME,"_delta_pairwise_average_distribution_per_feature.pdf", sep=""), width = 25, height = 25, units = "cm")
 
 
  # Plot de la distribución de delta corregida
 ggplot(get(DATAFRAME), aes(x=(delta_pairwise_corrected))) + 
    geom_histogram() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggsave (paste(wd_output,DATAFRAME,"_delta_pairwise_corrected_average_distribution.pdf", sep=""))
  # Plot de la distribución de delta para cada feature
 ggplot() + 
    aes(x = delta_pairwise_corrected) +
    geom_histogram() +
    facet_wrap(~feature, scales = "free")  +
    theme(axis.text.x = element_text(angle=90, hjust=1)) +
    ggsave (paste(wd_output,DATAFRAME,"_delta_pairwise_corrected_average_distribution_per_feature.pdf", sep=""), width = 25, height = 25, units = "cm")
 

 
}
```

## *Sanity: Delta vs initial diversity

```{r}
ggplot (c_ll_ki_n008_diversity_windows_c_ll_no_n008_diversity_windows, aes(x = log(ave_watterson.x), y = delta_watterson_corrected))+
  geom_point() +
  geom_smooth()

ggplot (c_ll_ki_n008_diversity_windows_c_ll_no_n008_diversity_windows, aes(x = log10(ave_pairwise.x), y = delta_pairwise_corrected))+
  geom_point() +
  geom_smooth ()

```




## *Delta summary

```{r}
DATAFRAMES=c("c_ll_ki_n008_diversity_windows_c_ll_no_n008_diversity_windows", "c_ll_ki_n008_diversity_windows_c_ll_po_n008_diversity_windows", "c_lp_sm_n019_diversity_windows_c_lp_do_n012_diversity_windows")

DELTA_MEASURES=c("delta_watterson", "delta_watterson_corrected", "delta_pairwise", "delta_pairwise_corrected")


for (DATAFRAME in DATAFRAMES) 
{
  for (DELTA_MEASURE in DELTA_MEASURES)
  {
  POP1=get(DATAFRAME) %>% select(pop.x) %>% mutate(pop.x=as.character(pop.x)) %>% .[1,1] 
  POP2=get(DATAFRAME) %>% select(pop.y) %>% mutate(pop.y=as.character(pop.y)) %>% .[1,1] 
  POPULATION1=get(DATAFRAME) %>% select(Population.x) %>% mutate(Population.x=as.character(Population.x)) %>% .[1,1] 
  POPULATION2=get(DATAFRAME) %>% select(Population.y) %>% mutate(Population.y=as.character(Population.y)) %>% .[1,1] 
  
ggplot(get(DATAFRAME), aes(x=feature, y=delta_watterson_corrected))+
      stat_summary(fun.data = "mean_cl_boot", size = 0.5) +
      coord_flip()
      theme(legend.position="top", axis.title.y=element_blank()) +
      scale_fill_manual(values=c("#588C7E", "#F2AE72", "#8C4646")) +
      scale_color_manual(values=c("#588C7E", "#F2AE72", "#8C4646")) +
  # PONER EL VALOR DEL GENERAL (ralla?)
#scale_fill_viridis_d(option = "magma")+
#scale_colour_viridis_d(option = "magma")+
      scale_x_discrete2(limits = rev)+

ggsave (paste(wd_output,DELTA_MEASURE, ".pdf", sep=""), width = 80, height = 80, units = "mm")
}
}

```


## *Sanity: Diversity one pop against other

```{r}
ggplot (c_ll_ki_n008_diversity_windows_c_ll_no_n008_diversity_windows, aes(x = ave_watterson.x, y = ave_watterson.y))+
  geom_point() +
  geom_smooth()
```

