---
title: "5.Admixture"
output: html_document
---
zcat c_ll_ba-c_ll_cr-c_ll_ka-c_ll_ki-c_ll_la-c_ll_no-c_ll_og-c_ll_po-c_ll_to-c_ll_tu-c_ll_vl-c_ll_ya-h_ll_ba.genolike.mafs.gz | tail -n+2 | wc -l
5437960


# Generating Admixture plot

## Converting our data in beagle format:


```{r, engine=bash, eval=FALSE}

THREADS=40                    # no. of computer cores used by bwa and samtools. 20 = OK, >20 = ask people first!

RUTA=/home/mlucena/ANGSD_analysis/depth_calculus 

read POP mean sd mean_truncated sd_truncated maxDepth minDepth maxDepth_truncated minDepth_truncated < $RUTA/"${POP}"_mean_sd_depthGlobal_lynx_per_pop_mean_folds_0.95.csv

ANGSD="/opt/angsd/angsd"
NGSTOOLS="/opt/angsd/angsd/misc"
REF="/home/GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/lp23_without_repetitive_transposable_low_complexity.fa"
ANC="/home/GRUPOS/grupolince/reference_genomes/lynx_rufus_genome/c_lr_zz_0001_recal1.fa"
FILTER1=" -uniqueOnly 1 -remove_bads 1 -only_proper_pairs 1 -baq 1 -C 50 "
FILTER2=" -minMapQ 30 -minQ 20 -doCounts 1 "
N_IND=$(echo ${POP: -3} )
MIN_IND=$(expr $N_IND / 2)
SNP_PVAL=$(expr 1 / ($N_IND * 2))
  
RUTA=/home/mlucena/ANGSD_analysis 

POP="c_ll_ba-c_ll_cr-c_ll_ka-c_ll_ki-c_ll_la-c_ll_no-c_ll_og-c_ll_po-c_ll_to-c_ll_tu-c_ll_vl-c_ll_ya-h_ll_ba"
POP_MODIF="x_ll_xx"

# Sanity checks:

echo $POP
echo $N_IND
echo $MIN_IND
echo $maxDepth_truncated
echo $minDepth_truncated

# Running GL:

$ANGSD/angsd -nThreads $THREADS --bam $RUTA/intergenic_analysis/"$POP"_n*.intergenic.bamlist  \
-out "$POP".genolike \
$FILTER1 \
$FILTER2 \
-GL 1  -doGlf 2 \ 
-doMajorMinor 1 -doMaf 3 -SNP_pval $SNP_PVAL -skipTriallelic 1 \
-minInd  $MIN_IND -setMaxDepth $maxDepth_truncated -setMinDepth $minDepth_truncated


# Parameters chosen:
# FILTER1=" -uniqueOnly 1 -remove_bads 1 -only_proper_pairs 1 -baq 1 -C 50 "
# FILTER2=" -minMapQ 30 -minQ 20 -doCounts 1 "
# GL 1 --> Use samtools
# doGlf 2 --> beagle likelihood file
# doMajorMinor 1 --> Infer major and minor from GL (other options would be: infer major and minor from allele counts, from a file, use reference as major, use ancestral as major)
# doMaf 3 --> frequency (fixed major and minor/fixed major,unknownminor): Both are assumed to be known. Allele freq i obtained using based on GL.  
# -doMajorMinor 1 -doMaf 3. Example for estimating the allele frequencies both while assuming known major and minor allele but also while taking the uncertaincy of the minor allele inference into account. The inference of the major and minor allele is done directly from the genotype likelihood
# -SNP_pval [float]
# The p-value used for calling snaps. see Allele_Frequency_estimation for additional options --> I choose at least 1 ind heterozygote. 


# for i in $POP*
# do
# echo $i
# scp $i ${i/$POP/$POP_MODIF} 
# done

# Imputation in beagle (optional):

# java -jar /home/mlucena/software/beagle.jar like="$POP_MODIF".genolike.beagle.gz out=

# out se deja sin nada pq por defecto pone cosas
```

Population format might be a problem as beagle does not admit long names! 
Check if it is beagle, cause then is not a problem. 

## Running admixture

```{r, engine=bash, eval=FALSE}

MINMAF=0.05


for i in `seq 1 12`;
do
echo $i
/home/mlucena/software/NGSadmix -likes "$POP_MODIF".genolike.beagle.gz -K "$i" -P 20 -o "$POP_MODIF"_K"$i" -minMaf $MINMAF 
done    

```


using only SNP with af MAF above 5% (-minMaf 0.05).


./NGSadmix
Arguments:
	-likes Beagle likelihood filename
	-K Number of ancestral populations
Optional:
	-fname Ancestral population frequencies
	-qname Admixture proportions
	-o Prefix for output files
	-printInfo print ID and mean maf for the SNPs that were analysed
Setup:
	-seed Seed for initial guess in EM
	-P Number of threads
	-method If 0 no acceleration of EM algorithm
	-misTol Tolerance for considering site as missing
Stop chriteria:
	-tolLike50 Loglikelihood difference in 50 iterations
	-tol Tolerance for convergence
	-dymBound Use dymamic boundaries (1: yes (default) 0: no)
	-maxiter Maximum number of EM iterations
Filtering
	-minMaf Minimum minor allele frequency
	-minLrt Minimum likelihood ratio value for maf>0
	-minInd Minumum number of informative individuals
	
	
## Plotting results

First we need a file with our population information:

```{bash}

cat /home/mlucena/ANGSD_analysis/intergenic_analysis/$POP*.bamlist | rev | cut -d "/"  -f 1 | cut -d "_" -f 3,4,5 | rev  > $POP_MODIF.column1.borrar
cat /home/mlucena/ANGSD_analysis/intergenic_analysis/$POP*.bamlist | rev | cut -d "/"  -f 1 | cut -d "_" -f 2,3,4,5 | rev  > $POP_MODIF.column2.borrar

paste $POP_MODIF.column1.borrar $POP_MODIF.column2.borrar > $POP_MODIF.info

rm *borrar
```


### Creating a plot

Guardo este script para lanzarlo en secuencial para cada una de las ks:

```{bash}
nano Admix_plot.R
```

Script que guardo

```{r}
library(RColorBrewer)
args <- commandArgs(TRUE)
print(args)
admix<-t(as.matrix(read.table(args[1])))
pop<-read.table(args[2],as.is=T)

names(pop) <- c("pop", "ind")
pop$pop <- as.factor (pop$pop)
pop$pop <- factor (pop$pop, levels=c("c_ll_no", "h_ll_ba", "c_ll_cr",  "c_ll_po", "c_ll_la", "c_ll_ki",  "c_ll_vl", "c_ll_ya", "c_ll_tu", "c_ll_og", "c_ll_ka", "c_ll_to"))

admix<-admix[,order(pop[,1])]
pop<-pop[order(pop[,1]),]
n <- args[3]
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
pie(rep(1,n), col=sample(col_vector, n))
#setEPS()
# postscript(paste0(args[4],"_K",args[3],".eps"))

pdf(paste0(args[4],"_K",args[3],".pdf"), height=10, width = 30)
barplot(admix,col=col_vector,space=0,xlab="Individuals",ylab="Admixture")
text(tapply(1:nrow(pop),pop[,1],mean),-0.1,unique(pop[,1]),xpd=T,srt=60)
dev.off()

```

Lo lanzo de manera secuencial.

```{bash}
for i in `seq 1 12`;
do
echo $i
Rscript Admix_plot.R "$POP_MODIF"_K"$i".qopt "$POP_MODIF".info $i $POP_MODIF
done
```


Test para hacer pruebinas con k=2. 

```{r}

library(RColorBrewer)

admix<-t(as.matrix(read.table("/Users/marialucenaperez/ownCloud/publico/Eurasian_lynx_phylogeography/lynx_lynx_ADMIXTURE/x_ll_xx_K7.qopt"))) 
pop<-read.table("/Users/marialucenaperez/ownCloud/publico/Eurasian_lynx_phylogeography/lynx_lynx_ADMIXTURE/x_ll_xx.info",as.is=T)

names(pop) <- c("pop", "ind")
pop$pop <- as.factor (pop$pop)
pop$pop <- factor (pop$pop, levels=c("c_ll_no", "h_ll_ba", "c_ll_cr",  "c_ll_po", "c_ll_la", "c_ll_ki", "c_ll_vl", "c_ll_ya", "c_ll_tu", "c_ll_og", "c_ll_ka", "c_ll_to"))

admix<-admix[,order(pop[,1])]
pop<-pop[order(pop[,1]),]

n <- 7
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
pie(rep(1,n), col=sample(col_vector, n))

h<-barplot(admix,space=0,xlab="Individuals",ylab="Admixture")
text(tapply(1:nrow(pop),pop[,1],mean),-0.1,unique(pop[,1]),xpd=T,srt=60)
barplot(admix,col=col_vector,space=0,xlab="Individuals",ylab="Admixture")
text(tapply(1:nrow(pop),pop[,1],mean),-0.1,unique(pop[,1]),xpd=T,srt=60)

```



## Test solo con la region de Elena porque sospecho que ha pasado algo durante el proceso de creación del anterior archivo. 

```{bash}

# Test
cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_ADMIXTURE/test

REGIONFILE="/home/mlucena/ANGSD_analysis/depth_calculus/no_genes_Lypa_10000longest_center_final_slop20_dot.rf"
POP="c_ll_ba-c_ll_cr-c_ll_ka-c_ll_ki-c_ll_la-c_ll_no-c_ll_og-c_ll_po-c_ll_to-c_ll_tu-c_ll_vl-c_ll_ya-h_ll_ba"
ANGSD=/opt/angsd/angsd


$ANGSD/angsd -GL 1 -out "$POP".TEST.genolike -nThreads 40 -doGlf 2 -doMajorMinor 1 -SNP_pval 1e-6 -doMaf 1  -rf $REGIONFILE  -bam /home/mlucena/ANGSD_analysis/whole_genome_analysis/$POP*.bamlist

java -jar /home/mlucena/software/beagle.jar like=lynx_lynx.TEST.genolike.beagle.gz out=lynx_lynx.TEST.beagleOutput

```


Command line: java -Xmx27305m -jar beagle.jar
  like=lynx_lynx.TEST.genolike.beagle.gz
  out=lynx_lynx.TEST.beagleOutput

number of markers    =    4718
number of haplotypes =     148 (lynx_lynx.TEST.genolike.beagle.gz)

Phasing: iteration 1

Phasing: iteration 2

Phasing: iteration 3

Phasing: iteration 4

Phasing: iteration 5

Phasing: iteration 6

Phasing: iteration 7

Phasing: iteration 8

Phasing: iteration 9

Phasing: iteration 10


```{bash}

/home/mlucena/software/NGSadmix -likes lynx_lynx.TEST.genolike.beagle.gz -K 1 -P 40 -o "lynx_lynx.TEST" -minMaf 0.05 

```






