---
title: "5.Admixture"
output: html_document
---

# Generating Admixture plot

## Converting our data in beagle format:


```{r, engine=bash, eval=FALSE}

ANGSD=/opt/angsd/angsd
RUTA=/home/mlucena/ANGSD_analysis 
POP="c_ll_ba-c_ll_cr-c_ll_ka-c_ll_ki-c_ll_la-c_ll_no-c_ll_og-c_ll_po-c_ll_to-c_ll_tu-c_ll_vl-c_ll_ya-h_ll_ba"


$ANGSD/angsd -GL 1 -out "$POP".genolike -nThreads 40 -doGlf 2 -doMajorMinor 1 -SNP_pval 1e-6 -doMaf 1  -bam $RUTA/intergenic_analysis/"$POP"_n*.intergenic.bamlist

java -jar /opt/beagle/beagle.21Jan17.6cc.jar gl="$POP".genolike.beagle.gz out="$POP".beagleOutput

java -jar /opt/beagle/beagle.21Jan17.6cc.jar like="$POP".genolike.beagle.gz out="$POP".beagleOutput

java -jar /opt/beagle/beagle.21Jan17.6cc.jar gl=test.beagle.gz out=test.beagleOutput


java -jar /home/mlucena/software/beagle.jar like=c_ll_ba-c_ll_cr-c_ll_ka-c_ll_ki-c_ll_la-c_ll_no-c_ll_og-c_ll_po-c_ll_to-c_ll_tu-c_ll_vl-c_ll_ya-h_ll_ba.genolike.beagle.gz out=c_ll_ba-c_ll_cr-c_ll_ka-c_ll_ki-c_ll_la-c_ll_no-c_ll_og-c_ll_po-c_ll_to-c_ll_tu-c_ll_vl-c_ll_ya-h_ll_ba.beagleOutput
```

## Test solo con la region de Elena porque sospecho que ha pasado algo durante el proceso de creación del anterior archivo. 

```{bash}

# Test
cd /home/mlucena/ANGSD_analysis/intergenic_analysis/intergenic_ADMIXTURE/test

REGIONFILE="/home/mlucena/ANGSD_analysis/depth_calculus/no_genes_Lypa_10000longest_center_final_slop20_dot.rf"
POP="c_ll_ba-c_ll_cr-c_ll_ka-c_ll_ki-c_ll_la-c_ll_no-c_ll_og-c_ll_po-c_ll_to-c_ll_tu-c_ll_vl-c_ll_ya-h_ll_ba"
ANGSD=/opt/angsd/angsd


$ANGSD/angsd -GL 1 -out "$POP".TEST.genolike -nThreads 40 -doGlf 2 -doMajorMinor 1 -SNP_pval 1e-6 -doMaf 1  -rf $REGIONFILE  -bam /home/mlucena/ANGSD_analysis/whole_genome_analysis/$POP*.bamlist

```


## Running admixture


./NGSadmix
Arguments:
	-likes Beagle likelihood filename
	-K Number of ancestral populations
Optional:
	-fname Ancestral population frequencies
	-qname Admixture proportions
	-o Prefix for output files
	-printInfo print ID and mean maf for the SNPs that were analysed
Setup:
	-seed Seed for initial guess in EM
	-P Number of threads
	-method If 0 no acceleration of EM algorithm
	-misTol Tolerance for considering site as missing
Stop chriteria:
	-tolLike50 Loglikelihood difference in 50 iterations
	-tol Tolerance for convergence
	-dymBound Use dymamic boundaries (1: yes (default) 0: no)
	-maxiter Maximum number of EM iterations
Filtering
	-minMaf Minimum minor allele frequency
	-minLrt Minimum likelihood ratio value for maf>0
	-minInd Minumum number of informative individuals
	
	

```{r, engine=bash, eval=FALSE}


/home/mlucena/software/NGSadmix -likes "$POP".beagleOutput.gz -K 1 -P 40 -o "$POP" -minMaf 0.05 -printInfo
/home/mlucena/software/NGSadmix -likes c_ll_ba-c_ll_cr-c_ll_ka-c_ll_ki-c_ll_la-c_ll_no-c_ll_og-c_ll_po-c_ll_to-c_ll_tu-c_ll_vl-c_ll_ya-h_ll_ba.genolike.beagle.gz -K 1 -P 40 -o c_ll_ba-c_ll_cr-c_ll_ka-c_ll_ki-c_ll_la-c_ll_no-c_ll_og-c_ll_po-c_ll_to-c_ll_tu-c_ll_vl-c_ll_ya-h_ll_ba -minMaf 0.05 -printInfo
  
  ```